vC Language Reference Manual

### 2020-11-21 v2/libAhirV2/docs/vC_LRM.tex
Bug fix in single-source phi (removed special casing) and in llvm2aa block chain guard handling.

A \vC system consists of modules and memory spaces.
Each module may itself contain memory spaces.  Further,
each module has a {\em control-path} and a {\em data-path}.
The control-path is a Petri-net constructed using
a set of rules (as describe in \cite{ref:SameerPhD}).1
The data-path is essentially a directed graph of {\em operator } nodes
communicating with each other and the outside world using {\em wires}.

@@

The control-path is a Petri-net which consists of
places and transitions.  In \vC this Petri-net
is described hierarchically in terms of 
control-path elements. 
\begin{itemize}
\item A single transition is a control-path element. It
fires when all of its predecessors have (released) a token.
After it fires, a token is released to each of
its successors.
\begin{verbatim}
$T [fire]
\end{verbatim}
This defines a transition with name $fire$. 
\item A single place is a control-path element.  When
any of its predecessors releases a token, the place
takes it and forwards it to one of its successors.
\begin{verbatim}
$P [loopback]
\end{verbatim}
\item A {\em series region} is a sequence of control
path elements.  The predecessor of each element is
the previous element and the successor is the next
element in the sequence.  
The predecessor of
the first element is an implicit entry transition
of the series region, and the successor of the last
element is an implicit exit transition of
the series region.
\begin{verbatim}
;;[chain] {
       $T [req]
       $T [ack]
}
\end{verbatim}
\item A {\em parallel region} is a sequence of control
path elements.  The implicit entry transition
is the predecessor of all the elements, while the
implicit exit transition is the successor of
all the elements.
\begin{verbatim}
||[pardo] {
       ;;[job_1] { ... }
       ;;[job_2] { ... }
}
\end{verbatim}
\item A {\em branch region} is a sequence of control
path elements with additional statements which describe
the connections between this sequence (branches, merges).
\begin{verbatim}
<>[branch_example] {
       $P [Enter]
       Enter <-| ($entry)
       Enter |-> (job_1)
       ;;[job_1] { ... }
       ;;[job_2] { ... }
       $P [Chk]
       $P [End]
       Chk <-| (job_1)
       Chk |-> (job_1 job_2)
       End <-| (job_2)
       End |-> ($exit)
}
\end{verbatim}
In this example, the keywords \$entry and \$exit correspond
to implicit entry and exit transitions associated with the branch block.
The $<-|$ specifies that a token leaving the right will enter
the left (place) and $|->$ signifies that a token entering the left
(place) will enter one of the elements in the list on the right.
\item A {\em fork region} is a sequence of control path
elements with additional statements which describe 
forking and joining.
\begin{verbatim}
::[fork_example] {
       $entry &-> (job_1 job_2)
       ;;[job_1] { ... }
       ;;[job_2] { ... }
       ;;[job_3] { ... }
       $T [f12]
       f12 <-& (job_1 job_2)
       f12 &-> (job_3)
       $exit <-& (job_3)
}
\end{verbatim}
In this example, the keywords \$entry and \$exit correspond
to implicit transitions associated with the branch block.
The <-\& specifies that after a token leaves from every
element in the right list, a token will enter
the left (transition) and \&-> signifies that a token leaving the left
(transition) will be replicated so that a
token enters each of the elements in the list on the right.
\item A pipelined-fork block defines a fork block which can support
several active instantiations at the same time.
\begin{verbatim}
:|: [<name>] {
   same-structure-as-fork
   but-can-have-marked-joins
}
\end{verbatim}
This uses a special join called a marked-join
\begin{verbatim}
jT o<- (s1 1 s2 1)
(s1 can be $entry)
\end{verbatim}
Which says that jT is a join of s1 and s2, with a pre-condition
that s1 is assumed to have fired twice and s2 is assumed to have fired
twice before time began.  As a result it is possible for jT to fire
twice before s1 and s2 fire in normal operation. 
\end{itemize}

@@

>>>>

A \vC system consists of modules and memory spaces.
Each module may itself contain memory spaces.  Further,
each module has a {\em control-path} and a {\em data-path}.
The control-path is a Petri-net constructed using
a set of rules (as described in \cite{ref:SameerPhD}).
The data-path is essentially a directed graph of {\em operator } nodes
communicating with each other and the outside world using {\em wires}.

@@

The control-path is a Petri-net which consists of
places and transitions.  In \vC this Petri-net
is described hierarchically in terms of 
control-path elements. 
\begin{itemize}
\item A single transition is a control-path element. It
fires when all of its predecessors have (released) a token.
After it fires, a token is released to each of
its successors.
\begin{verbatim}
$T [fire]
\end{verbatim}
This defines a transition with name $fire$. 
\item A single place is a control-path element.  When
any of its predecessors releases a token, the place
takes it and forwards it to one of its successors.
\begin{verbatim}
$P [loopback]
\end{verbatim}
\item A {\em series region} is a sequence of control
path elements.  The predecessor of each element is
the previous element and the successor is the next
element in the sequence.  
The predecessor of
the first element is an implicit entry transition
of the series region, and the successor of the last
element is an implicit exit transition of
the series region.
\begin{verbatim}
;;[chain] {
       $T [req]
       $T [ack]
}
\end{verbatim}
\item A {\em parallel region} is a sequence of control
path elements.  The implicit entry transition
is the predecessor of all the elements, while the
implicit exit transition is the successor of
all the elements.
\begin{verbatim}
||[pardo] {
       ;;[job_1] { ... }
       ;;[job_2] { ... }
}
\end{verbatim}
\item A {\em branch region} is a sequence of control
path elements with additional statements which describe
the connections between this sequence (branches, merges).
\begin{verbatim}
<>[branch_example] {
       $P [Enter]
       Enter <-| ($entry)
       Enter |-> (job_1)
       ;;[job_1] { ... }
       ;;[job_2] { ... }
       $P [Chk]
       $P [End]
       Chk <-| (job_1)
       Chk |-> (job_1 job_2)
       End <-| (job_2)
       End |-> ($exit)
}
\end{verbatim}
In this example, the keywords \$entry and \$exit correspond
to implicit entry and exit transitions associated with the branch block.
The 
\begin{verbatim}
<-|
\end{verbatim}
 specifies that a token leaving the right will enter
the left (place) and 
\begin{verbatim}
|->
\end{verbatim}
signifies that a token entering the left
(place) will enter one of the elements in the list on the right.
\item A {\em fork region} is a sequence of control path
elements with additional statements which describe 
forking and joining.
\begin{verbatim}
::[fork_example] {
       $entry &-> (job_1 job_2)
       ;;[job_1] { ... }
       ;;[job_2] { ... }
       ;;[job_3] { ... }
       $T [f12]
       f12 <-& (job_1 job_2)
       f12 &-> (job_3)
       $exit <-& (job_3)
}
\end{verbatim}

In this example, the keywords \$entry and \$exit correspond
to implicit transitions associated with the branch block.
The
\begin{verbatim}
 <-&
\end{verbatim}
 specifies that after a token leaves from every
element in the right list, a token will enter
the left (transition) and the
\begin{verbatim}
&->
\end{verbatim}
signifies that a token leaving the left
(transition) will be replicated so that a
token enters each of the elements in the list on the right.
\item A pipelined-fork block defines a fork block which can support
several active instantiations at the same time.
\begin{verbatim}
:|: [<name>] {
   same-structure-as-fork
   but-can-have-marked-joins
}
\end{verbatim}
This uses a special join called a marked-join
\begin{verbatim}
jT o<- (s1 1 s2 1)
(s1 can be $entry)
\end{verbatim}
Which says that jT is a join of s1 and s2, with a pre-condition
that s1 is assumed to have fired and s2 is assumed to have fired
before time began.  As a result it is possible for jT to fire
before s1 and s2 fire in normal operation. 
\end{itemize}

@@

\subsection{Special constructs in the control path}

Some special constructs are used in the control path.
\begin{itemize}
\item Transition merge:  This is specified as
\begin{verbatim}
$transition_merge [merge_name] (p q r) (s)
\end{verbatim}
Where p, q, r are transition, and s fires whenever any 
of p,q,r fire.
\item  Phi-sequencer:   This is specified
as
\begin{verbatim}
$phisequencer [phiseq_name] :  
    trigger_1
       src_1_sample_req src_1_sample_ack
       src_1_update_req src_1_update_ack
    trigger_2
       src_2_sample_req src_2_sample_ack
       src_2_update_req src_2_update_ack:
    phi_sample_start phi_sample_completed
       phi_update_start phi_update_completed :
    trigger_1_sample_req  trigger_2_sample_req:
       phi_mux_ack
\end{verbatim}
and it describes the control connections between the Phi operator and
its sources.
\item Loop terminator: This is specified as
\begin{verbatim}
$terminate (loop_exit loop_taken loop_body_done) 
              (loop_back loop_terminate)
\end{verbatim}
The loop terminator looks at the result of the do-while condition
check via the input transitions loop\_exit and loop\_taken, 
and uses the loop\_body\_done transition to determine the
number of transitions in flight and hence to trigger the
next do-while iteration (firing of loop\_back) or to
decide that the loop has terminated.
\end{itemize}
